cardinalR: Generating Interesting High-Dimensional Data Structures

### 2025-12-19 paper-cardinalR-submission-v1/paper-cardinalR.tex
SUBMITTED TO R journal !!!!

{\centering \includegraphics[width=0.8\linewidth,alt={The figure is a workflow diagram showing the process for generating high-dimensional clustered data. Boxes or nodes represent steps such as specifying input parameters (number of points, number of clusters, cluster locations, shapes, scaling, rotations, and optional background noise), generating each cluster with a shape generator, optionally rotating or scaling clusters, combining them into a single dataset, adding background noise if desired, and labeling each observation by cluster shape. Arrows indicate the flow of the process from parameter specification to the final labeled dataset. The diagram uses a clear layout with distinct steps, showing how data are constructed in sequence.}]{figures/cardinalR_workflow}

@@

The source material for this paper is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}. This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template. These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.

>>>>

{\centering \includegraphics[width=0.8\linewidth,alt={The figure is a workflow diagram showing the process for generating high-dimensional clustered data. Boxes or nodes represent steps such as specifying input parameters (number of points, number of clusters, cluster locations, shapes, scaling, rotations, and optional background noise), generating each cluster with a shape generator, optionally rotating or scaling clusters, combining them into a single dataset, adding background noise if desired, and labeling each observation by cluster shape. Arrows indicate the flow of the process from parameter specification to the final labeled dataset. The diagram uses a clear layout with distinct steps, showing how data are constructed in sequence.}]{../../../../../../private/var/folders/4d/q_zhr19n0z7d8hx3xblybm8rjpxby3/T/RtmpRmAUiT/file794b45c38a65}

@@

The source material for this paper, including the full datasets and figures, is available at \href{https://github.com/JayaniLakshika/paper-cardinalR}{github.com/JayaniLakshika/paper-cardinalR}. This article is created using \CRANpkg{knitr} \citep{yihui2015} and \CRANpkg{rmarkdown} \citep{yihui2018} in R with the \texttt{rjtools::rjournal\_article} template. These \texttt{R} packages were used for this work: \texttt{cli} \citep{gabor2025}, \texttt{tibble} \citep{kirill2023}, \texttt{gtools} \citep{gregory2023}, \texttt{dplyr} \citep{hadley2023}, \texttt{stats} \citep{core2025}, \texttt{tidyr} \citep{hadley2024}, \texttt{purrr} \citep{hadley2025}, \texttt{mvtnorm} \citep{alan2009}, \texttt{geozoo} \citep{barret2016}, and \texttt{MASS} \citep{venables2002}.
